Using best practices for forecasting biodiversity changes in breeding birds

### 2017-07-07 manuscript/draft.Rmd
Edits to Methods and Results sections
* Expand methods into full paragraph draft and add refs
* Add recommend results changes from #118
* General editorial work on results

* BBS + retriever
* Predictor variables
  * Monthly climate data from interpolated weather stations (PRISM), converted
    to BIOCLIM variables. Many of these variables were tightly correlated, so
    we only included the 8 variables selected by Harris's (2015) BBS analysis
  * Weather/climate and NDVI were each calculated separately for each year,
    rather than averaged over the full multi-decade period (as Worldclim does)
  * NDVI (GIMMS)
  * Altitude (SRTM 90m Digital Elevation Database via raster::getData)
* Data summary: 
  * Number of sites, species, unique observers
  * central moments of richness (mean about 51, sd about 12 spp)
  * The within-site variation was much smaller (sd about 5 spp), making the
    site-level distributions underdispersed.  
    * This underdispersion implies
      that common count models such as binomial, Poisson, negative binomial 
      won't fit well.
    * For this reason (and because the time series methods assumed 
      Gaussian errors), we defined model performance throughout this paper in 
      terms of Gaussians (e.g. mean squared error), rather than with count
      distributions.
        * Some of the modeling approaches (SDMs and JSDMs) did use discrete
          distributions internally, however.
* Splitting the data temporally
  * Our data set began in `r settings$start_yr`, which was the first year that 
    NDVI data was available from GIMMS
  * All the models were trained on the data that was collected through 
    `r settings$last_train_year`
  * Sites were included in the analysis if there were at least 
    `r settings$min_num_yrs` years of data available during this period.
  * The models were tested/evaluated on the data from 
    `r settings$last_train_year + 1` `r settings$end_yr`.

  Bird species richness was compiled from the Breeding Bird Survey (BBS) [@pardieck2017] obtained using the Data Retriever software [@morris203]. The BBS data was initially filtered to exclude all nocturnal, cepuscular, and aquatic species (since these species are not well sampled by BBS methods), as well as unidentified species, and hybrids. All data from routes that did not meet BBS acceptable criteria were excluded. Only routes within the continental U.S.A were used so as to match the climatic dataset.  
  We used the 4 km resolution PRISM data [@prism] to calculate eight bioclimatic variables identified as relevant to bird distributions [@harris2015] for each year in the dataset: mean diurnal range, isothermality, max temperature of the warmest month, mean temperature of the wettest quarter, mean temperature of the driest quarter, precipitation seasonality, precipitation of the wettest quarter, precipitation of the warmest quarter. Satellite derived NDVI with an 8 km resolution was from the NDIV3g dataset [@pinzon2014] and was available from the time period 1981-2013. Elevation was from the SRTM 90m elevation dataset [@jarvis2008] obtained using the R package raster [@hijmans2014]. BBS route starting point coordinates were used to extract all environmental data. Climatic variables and NDVI were aggregated to the average values within a 40 km radius of the start point, equal to the length of a BBS route. NDVI was also subset to average summer (May, June, July) and winter (December, January, Feburary) values.
  Data from 1982 (the first complete year of NDVI data) to 2003 were used for training models and data from 2004 to 2013 was used for testing. Only BBS routes with at least 25 yearly observations in the years 1982-2015 were used.

### Accounting for site \& observer effects
* Observer effects are inherent in large data sets taken by different 
  observers, and are known to occur in BBS [citation].
* While these biases are known to exist, they are difficult to quantify: even if
  the different observers ran the same transects under identical conditions (which 
  they don't), we would still only have an estimate of their differences, not an 
  exact value.
* We estimated the observer effects, along with our uncertainty about those
  effects' size & direction, with a linear mixed model.
* This model partitions the variance in observed richness values into site-level
  variance, observer-level variance, and residual variance (e.g. variation within
  a site from year to year or from morning to morning).
  * No fixed effects in this model (e.g. environmental predictor variables because 
    we're not interested in explaining differences among sites at this point; we 
    just want to account for their existence while estimating observer effects.
  * Subtracting off the estimated observer effect provides an estimate of how many
    would have been found by a "typical" observer on that day; in this sense, it 
    enables us to correct for observer differences. We then used this "corrected" 
    version of the data for our time-series models such as the AR1, etc.
* A simple model like this one can't tell us exactly how each observer differed from
  the others, so it is important to represent our uncertainty about these differences.
  Here, we did so by collecting [[N]] Monte Carlo samples from the model's posterior
  distribution, using Stan. This provided us with [[N]] plausible values for each 
  observer, so we did not have to commit to a single value.
* While some methods for predicting richness might be able to use the uncertain
  estimates of the observer effects directly, we take the simpler approach of just
  fitting each of the downstream models with each of the Monte Carlo samples. This
  increases the amount of computational work by a factor of [[N]], but CPU time
  wasn't our limiting resource.

@@

[[note that these all use a Gaussian error distribution]]

**Baseline models.** The two simplest models we fit treated site-level richness
observations either as uncorrelated noise around a site-level constant (the 
"average" model) or as an autoregressive model with a single year of history 
(the "naive" model). The predictions from these two models are qualitatively 
different: predictions from the "average" model are centered on the average 
richness and the confidence intervals are narrow and constant-width. The
"naive" model, in contrast, predicts that future observations will be similar to
the final observed value (e.g. the value observed in `$r settings$last_train_year`), 
and the confidence intervals expand rapidly as the predictions extend farther into 
the future.  Each of these model types was fit separately to each BBS route in two
ways: first, we fit the model to the raw richness values, without accounting for
possible differences among observers at the site.  Then, we fit the models to the
residuals from our observer model (i.e. fitting the model to an estimate of how
many species a typical observer would have found, rather than to the raw data). 
In the latter case, we averaged across [[N]] models that were fit separately to 
the [[N]] Monte Carlo estimates of the observer effects, to account for our
uncertainty in the true values of those effects.

**Time series models.** Like the two baseline models, these models only looked at 
one site at a time, but they could include additional complexity. As above, these
models were trained both with the raw richness time series and with Monte Carlo
estimates of the observer-corrected time series. The  Auto-ARIMA models (based on 
Hyndman's 2016 `auto.arima` function) could include an autoregressive component 
(as in the "naive" model, but with the possibility of longer-term dependencies
in the underlying process), a moving average component (where the noise can have 
serial autocorrelation) and an integration/differencing component (so that the 
analysis could be performed on sequential differences of the raw data, rather than
on the observed time series itself). The `auto.arima` function chooses whether to 
include each of these components, and how many terms to include for each one, using
AICc, as described in Hyndman (2016). We used the default settings for this function,
apart from setting `seasonal = FALSE` because our time series does not include a 
seasonal component.  [[Note: should probably record somewhere what order model
the auto.arima function tends to choose.  Eyeballing the results, seems to be most 
similar to the naive model, but would be good to check.]] We also fit an 
"Auto-ARIMA + environment" model, in which each site-level time series could 
also be predicted by annual changes in our environmental variables at that site.

@@

So far, each model has been fit to one site at a time, with no information 
(apart from observer effects) shared among sites. When continental-scale data 
is available, ecologists often use correlative models to predict richness 
changes over time from environmental data. Here, we tested three variants of 
this approach. Here, we did *not* include the site-level random effects as 
predictors, meaning that this approach implicitly assumes that two sites 
with identical Bioclim, elevation, and NDVI values should have identical 
richness distributions. As above, we included observer effects & the 
associated uncertainty by running this model [[N]] times (once per MCMC 
sample).

**"Macroecological" model.** The most straightforward way to predict richness is 
to fit a richness model. We chose the `gbm` package for fitting boosted 
regression trees as our example of such a model. Boosted regression trees is a 
common approach in the species distribution modeling literature [refs] and 
works by fitting thousands of small tree-structured models sequentially, with
each tree optimized to reduce the error of its predecessors. For the reasons 
discussed above, this model was optimized using a Gaussian likelihood. [[other 
gbm settings go here.]] Compared to SDMs, this lets us predict richness directly 
instead of using hundreds of noisy species-level estimates, but it also discards 
any potentially-useful information regarding species turnover.

**SDMs.**
* Probably the most common approach?
* These models say that species' occurrence probabilities depend only on the values of
  the environmental variables at the time and place of the transect, not on the history
  of the environment or the history of the birds at that site
* Predict each species' occurrence probabilities individually, then add up the predictions
* [[Something about not using thresholds that throw away information]]
* Taking this approach in the face of uncertainty about observer effects would either 
  require a customized model (cite "neighborly advice" paper), or fitting [[N]] models
  times the number of species.
* To make it feasible to fit this large number of models, we used a custom wrapper 
  around the randomForest package. Random forests are flexible models that are built 
  by combining many independent sub-models ("trees") fit to different versions of the training
  data, which made it straightforward to adapt them to our situation: we simply fit one 
  tree using each of the Monte Carlo estimates of our observer effects and averaged the 
  predictions of the resulting forest.
* For the uncertainty, we used Gaussian approximations, based on the means and variances
  of the sums of independent Bernoulli random variables (cite Calabrese et al.).

**JSDM** 
* Joint species distribution model (JSDM) is a new approach that makes predictions
  about the full community composition, rather than modeling it species-by-species.
* In stacked single-species SDMs, the species-level occurrences are treated as 
  independent coin-flips whose proability depends only on the environment.
* JSDMs remove this independence assumption, and explicitly account for the
  possibility that a site will be much more (or less) suitable for birds in 
  general than one would expect based on a few environmental measurements alone.
* As a result, JSDMs do a better job of representing our uncertainty about 
  richness, while stacked SDMs underestimate it (Harris 2015).
* We used the `mistnet` package (Harris 2015) because it is the only JSDM that
  describes species' environmental associations with nonlinear functions.
* The mistnet package doesn't "know" about time series or spatial autocorrelation,
  so it did not share information among the repeated runs of the same transect.

### Ensembles
* Ensembles. Averaging predictions from multiple models tends to reduce the 
  noise in the estimates, and can thus lead to better predictions. Choosing the
  weights for the ensemble isn't straightforward because our estimates of model 
  error on the training set are biased in favor of models that
  overfit. [[Another option is to spatially cross-validate on the training set]]

@@

* Evaluated each year between  `r settings$last_train_year` and `r settings$end_yr`
* When a model was run multiple times (e.g. once per MCMC sample for the observer
  estimates), we used the mean of those runs' point estimates as our final point 
  estimate and we calculated the uncertainty using the law of total variance (i.e.
  the average of the model runs' variance, plus the variance in the point estimates).
* Metrics:
  * (R)MSE 
    * How far, on average, are the models' predictions from the true value?
    * [[RMSE has better units than MSE (species versus squared species), but MSE is
      nice because it's additive (i.e. MSE of 2 is twice as big as MSE of 1).]]
  * Calibration/bias
    * In general, or in specific situations, is the model consistently predicting
      that richness will be too high or too low?
    * Can also report this value in terms of RMSE or MSE
  * Coverage of 95% confidence intervals
  * Gaussian deviance (simultaneously rewards good point estimates, precision, 
    and coverage)

Looked at how error changed as the time horizon of forecasting lengthened
* Error will generally tend to increase
* Error might flatten out after some number of years
* Error might increase at different rates for different methods
  * One model might be better at short time scales \& worse at long time scales

@@

The mixed model (which formed the basis of our observer-corrected "average"
predictions) found that 70% of the variance in richness in the training set
could be explained by differences among sites, and 21% could be explained by
differences among observers. The remaining 9% represents residual variation,
where a given observer might report a different number of species on different
runs of the same transect. We found remarkably little autocorrelation in these
residuals ([[about 0.04, +/- something]]), suggesting richness is close to
stationary on these time scales, once observer differences are accounted for.

The substantial uncertainty associated with differences among observers caused
reliable improvements in both mean-squared-error and deviance when observer
effects were controlled for (Figure 4). This improvement was smallest for the
"Naive" model, as discussed below.

In general, each model's error increased more-or-less linearly with time,
although the slopes and initial error rates varied across models (Figure 3). The
site-level models (i.e. the baselines and the ARIMA methods) tended to do very
well (Figure 3), especially in terms of RMSE and especially on short time
horizons. For most combinations of year and metric, the best-performing model
was one of the two baselines. Usually, the "naive" model had the lowest error
when observer effects were not explicitly accounted for, and the "average" model
had the lowest error when they were. We discuss the reasons for this difference
in the next section.

Our two error metrics produced qualitatively similar results in most cases, with
the notable exception of the stacked SDMs (Figure 3A and 3B). Deviance, unlike
RMSE, depends on the full predictive distribution of a model, including the
model's estimated uncertainty. As noted elsewhere [@harris2015], these models
are generally far too confident, with observed richness values frequently
falling well outside their confidence limits (Figure 3C). As a result, the
stacked SDMs' deviance was dramatically higher than the next-worst model even
though its RMSE was not unusual. On the other hand, the richness-GBM model and
especially the mistnet JSDM had reasonably well-calibrated uncertainty
estimates, even in the presence of uncorrected observer biases (Figure 3C). This
prevented their deviance from rising as quickly as the site-level models when
the observer population changed (Figure 3B).

In general, the models produced confidence intervals that were too narrow,
indicating overconfident predictions. Of these, the random forest-based SDM
stack was the most overconfident, as discussed above. As one might expect,
correcting for observer differences improved the coverage of most models' 95%
intervals. Interestingly, this effect was very small for mistnet (its
coverage only improved from 92% to 93%), which seems to indicate that its
latent variables were already accounting for observer-related uncertainty. 
The "naive" model was the only model whose confidence intervals were too wide
(Figure 3), which can be attributed to the rapid rate at which
these intervals expand as the time lag increases (Figure 1).

>>>>

Bird species richness was obtained from the Breeding Bird Survey (BBS)
[@pardieck2017] using the Data Retriever software [@morris2013] and
rdataretriever R package [@mcglinn2017]. The BBS data was filtered to exclude
all nocturnal, cepuscular, and aquatic species
[since these species are not well sampled by BBS methods; @hurlbert2005], as
well as unidentified species, and hybrids. All data from routes that did not
meet BBS acceptable criteria were excluded. Only routes within the continental
United States were used so as to match the climate dataset. The resulting
dataset included [[N]] species and [[N]] sites. Among sites richness varied from
[[N]] to [[N]] with an average richness of [[N]] species. Within-site richness
variation (among years) was small (~5 species).

Environmental data included a combination of elevation, bioclimatic variables
and a remotely sensed vegetation index (the normalized difference vegetation
index; NDVI), all of which are know to influence richness and distribution in
the BBS data. We used the 4 km resolution PRISM data [@prism] to calculate eight
bioclimatic variables identified as relevant to bird distributions [@harris2015]
for each year in the dataset: mean diurnal range, isothermality, max temperature
of the warmest month, mean temperature of the wettest quarter, mean temperature
of the driest quarter, precipitation seasonality, precipitation of the wettest
quarter, precipitation of the warmest quarter. Satellite derived NDVI, a primary
correlated of richness in BBS data [@hurlbert2002], with an 8 km resolution was
obtained from the NDIV3g dataset [@pinzon2014] and was available from the time
period 1981-2013. Average summer (May, June, July) and winter (December,
January, Feburary) NDVI values were used as predictors. Elevation was from the
SRTM 90m elevation dataset [@jarvis2008] obtained using the R package raster
[@hijmans2016]. All environmental variables were aggregated to the average
values within a 40 km radius (the length of a BBS route) of the starting point
of each BBS route.

Data from `r settings$start_yr` (the first complete year of NDVI data) to `r
settings$last_train_year` were used for training models and data from `r
settings$last_train_year + 1` to `r settings$end_yr` was used for testing. Only
BBS routes with at least `r settings$min_num_yrs` annual observations in the
years 1982-2015 were used.

### Accounting for observer effects

Observer effects are inherent in large data sets taken by different observers,
and are known to occur in BBS [@sauer1994]. We estimated the observer effects,
and associated uncertainty about those effects, with a linear mixed model with
observer as a random effect. Because observer and site are strongly related (an
observer tends to repeatedly sample the same site), site was also included as a
random effect to ensure that inferred deviations were actually observer-related
(as opposed to being related to the sites that a given observer happened to
see). The resulting model partitions the variance in observed richness values
into site-level variance, observer-level variance, and residual variance
(e.g. variation within a site from year to year), and can therefore also be used
directly as the "average" baseline model (see below). The estimated observer
effects can be subtracted from the richness values for a particular observer to
provide an estimate of how many species would have been found by a "typical"
observer. To incorporate uncertainty in these "corrected" richness values into
the forecasting models we collected [[N]] Monte Carlo samples from the model's
posterior distribution, using Stan, and fit each of the downstream models with
each of the Monte Carlo samples.

@@

We fit three sets of models that were fit to each site independently. These
models were fit to each BBS route both on the raw richness values and on the
residuals from our observer model. When correcting for observer effects, we
averaged across [[N]] models that were fit separately to the [[N]] Monte Carlo
estimates of the observer effects, to account for our uncertainty in the true
values of those effects. All of these models use a Gaussian error distribution.

**Baseline models.** We used two simple models as baselines for comparison to
the fits of more complex models. These baselines treated site-level richness
observations either as uncorrelated noise around a site-level constant (the
"average" model) or as an autoregressive model with a single year of history
(the "naive" model) [@hyndman2014]. Predictions from the "average" model are
centered on the average richness and the confidence intervals are narrow and
constant-width. The "naive" model, in contrast, predicts that future
observations will be similar to the final observed value (e.g. the value
observed in `$r settings$last_train_year`), and the confidence intervals expand
rapidly as the predictions extend farther into the future.

**Time series models.** We used Auto-ARIMA models [based on the `auto.arima`
function in @hyndman2016] to represent a array of different time-series modeling
approaches. These models can include an autoregressive component (as in the
"naive" model, but with the possibility of longer-term dependencies in the
underlying process), a moving average component (where the noise can have serial
autocorrelation) and an integration/differencing component (so that the analysis
could be performed on sequential differences of the raw data, accommodating more
complex patterns including trends). The `auto.arima` function chooses whether to
include each of these components, and how many terms to include for each one,
using AICc [@hyndman2016]. Since there is no seasonable component to the BBS
time-series we did not include a season component in these models. Otherwise we
used the default settings for this function. [[Note: should probably record
somewhere what order model the auto.arima function tends to choose.  Eyeballing
the results, seems to be most similar to the naive model, but would be good to
check.]] We also fit an "Auto-ARIMA + environment" model, in which each
site-level time series could also be predicted by annual changes in our
environmental variables at that site.

@@

In contrast to the time-series models, most attempts to predict species richness
focusing on using correlative models to predict richness changes over time from
environmental data. We tested three common variants of this approach - direct
modeling of species richness, stacking individual species distribution models,
and joint species distribution models. Following the standard approach,
site-level random effects were not included in these models as predictors,
meaning that this approach implicitly assumes that two sites with identical
Bioclim, elevation, and NDVI values should have identical richness
distributions. As above, we included observer effects & the associated
uncertainty by running these models [[N]] times (once per MCMC sample).

**"Macroecological" model.** We used a generalized boosted regression tree model
using the `gbm` package [@ridgeway2017] to directly model species richness as a function of
environmental variables. Boosted regression trees are a form of tree based
modeling that work by fitting thousands of small tree-structured models
sequentially, with each tree optimized to reduce the error of its
predecessors. They are flexible models that are considered well suited for
prediction [@elith2008]. This model was optimized using a Gaussian
likelihood. [[other gbm settings go here.]]

**Species Distribution Models.** Species distribution models (SDMs) model
individual species' occurrence probabilities using environmental
variables. Species level models are used to predict richness by summing the
predicted probability of occupancy across all species at a site. We fit SDMs
using random forests, a common approach in the species distribution modeling
literature [[refs; would be good to provide a source that says these are a good
choice]]. Predicted probabilities of occurrence were summed directly to predict
the number of species. This avoids known problems with the use of thresholds for
determining whether or not a species will be present at a site [[refs]]. When
incorporating observer effects we avoided fitting [[N x S]] models ([[N]] MCMC
samples * [[S]] species) by fitting one regression tree using each of the Monte
Carlo estimates of our observer effects and averaging the predictions of the
resulting forest [[citation?]]. We used Gaussian approximations for uncertainty,
based on the means and variances of the sums of independent Bernoulli random
variables [[cite Calabrese et al.]].

**Joint Species Distribution Models.** Joint species distribution models (JSDMs)
are a new approach that makes predictions about the full composition of a
community instead of modeling each species independently. JSDMs remove the
assumed independence among species and explicitly account for the possibility
that a site will be much more (or less) suitable for birds in general, or
particular groups of birds, than one would expect based on the available
environmental measurements alone. As a result, JSDMs do a better job of
representing our uncertainty about richness, while stacked SDMs underestimate it
[@harris2015; [[other refs, e.g., Clark]]]. We used the `mistnet` package
[@harris2015] because it is the only JSDM that describes species' environmental
associations with nonlinear functions.

@@

We evaluated the performance of each model for forecasting using the data for
each year between `r settings$last_train_year` and `r settings$end_yr`. We used
three metrics for evaluating performance: 1) root-mean-square error (RMSE) to
determine how far, on average, the models' predictions were from the observed
value; 2) the 95% prediction interval coverage to determine how well the models
did at predicting uncertainty; and 3) deviance as an integrative measure of fit
incorporating good point estimates, precision, and coverage. When a model was
run multiple times for the purpose of correcting for observer effects, we used
the mean of those runs' point estimates as our final point estimate and we
calculated the uncertainty using the law of total variance (i.e.  the average of
the model runs' variance, plus the variance in the point estimates). In addition
to looking at forecast performance in general we evaluated how performance
changed as the time horizon of forecasting increased by plotting performance
metrics against the forecast horizon.

Variance in species richness within sites was lower than predicted by several
common count models (i.e. richness was underdispersed), so these count models
would have difficulty fitting the data. Since richness also had a relatively
large mean ([[N]]) and all models already produce continuous richness estimates
(due to summing probabilities in the (J)SDM case), we defined model performance
for all models in terms of continuous Gaussian errors instead of using discrete
count distributions.

@@

The site-observer model found that 70% of the variance in richness in the
training set could be explained by differences among sites, and 21% could be
explained by differences among observers. The remaining 9% represents residual
variation, where a given observer might report a different number of species in
different years. There was little temporal autocorrelation in these residuals
([[about 0.04, +/- something]]), suggesting that richness is close to stationary
on these time scales once observer differences are accounted for.

When comparing forecasts for richness across sites all methods performed well
(Figure 2; R^2 XX-XX%). However stacked and joint SDMs and the macroecological
model both failed to successfully forecast the highest richness sites resulting
in a noteable clustering of predicted values near ~60 species and the poorest
performance in the cross-site comparison (R^2 XX-XX%).

For most combinations of metric and forecast time lag, the best-performing model
for forecasting within site variability was either the "average" baseline model
or the ARIMA model (Figure 3). In general, each model's error increased
more-or-less linearly with time, although the slopes and initial error rates
varied across models (Figure 3). The site-level models (i.e. the baselines and
the ARIMA methods) tended to do very well (Figure 3) on short time horizons,
especially in terms of RMSE.

In general, the models produced confidence intervals that were too narrow,
indicating overconfident predictions. Of these, the random forest-based SDM
stack was the most overconfident. As noted elsewhere [@harris2015], these models
are generally far too confident, with observed richness values frequently
falling well outside their confidence limits (Figure 3C). The "naive" model was
the only model whose confidence intervals were too wide (Figure 3), which can be
attributed to the rapid rate at which these intervals expand as the time lag
increases (Figure 1). The JSDM and macroecological model has the best coverage.

RMSE and deviance produced qualitatively similar results in most cases, with the
notable exception of the stacked SDMs (Figure 3A and 3B). Deviance, unlike RMSE,
depends on the full predictive distribution of a model, including the model's
estimated uncertainty. As a result, the stacked SDMs' deviance was notably
higher than the next-worst model even though its RMSE was not unusual. On the
other hand, the richness-GBM model and especially the mistnet JSDM had
reasonably well-calibrated uncertainty estimates, even in the presence of
uncorrected observer biases (Figure 3C). This prevented their deviance from
rising as quickly as the site-level models when the observer population changed
and resulted in better comparative performance for these models using deviance
(Figure 3B).

Accounting for the uncertainty associated with differences among observers
resulted in improvements to all metrics of fit (Figure 4). The one exception was
RMSE of the "Naive" model, which remained unchanged. The improvements in point
estimates primarily resulted from large improvements in a small number of
forecasts, presumably ones where a large difference between observers resulted
in a large shift in predicted richness values that the models could not
otherwise accommodate. Most forecasts exhibited only small changes in
RMSE. However these changes were sufficient to reorder which models performed
best overall. For example, the "naive" model generally had the lowest average
error when observer effects were not explicitly accounted for, switching to the
average and ARIMA models once observer differences were corrected. Improvements
in coverage were notable for all models except the naive and JSDM, meaning that
incorporating observer effects resulted in improved quantification of
uncertainty. The lack of change in the mistnet JSDM seems to indicate that its
latent variables were already accounting for observer-related uncertainty. In
combination changes in point estimates and coverage resulted in changes in the
best models based on deviance at different time scales. The macroecological
model and the JSDM appeared to narrowly outperform the average and ARIMA models
when the forecast horizon was greater than five years when not correcting for
observers, but after the correction the average and ARIMA models maintained a
small average advantage out to at least ten years.
